Interpreting Random Forest Predictions for Firearm Identification Using LIME

### 2018-10-25 paper/paper.Rmd
added if else statements to paper

@@

# Join the two cleaned Hamby 224 sets into one testing set
hamby224_test <- suppressWarnings(bind_rows(hamby224_set1_cleaned,
                                            hamby224_set11_cleaned))

# Save the test data as a .csv file
write.csv(hamby224_test, "../data/hamby224_test.csv", row.names = FALSE)
```

@@

```{r echo = FALSE}
# Apply the lime function from the lime package (with a seed set)
# Note that the as_classifier must be added since rtrees is from the
# randomForest package and not fit using caret or one of the other 
# available models specified in the lime package. Additionally, the
# randomForest package must be loaded in order to run the functions
# from the lime package. (I should check on exactly how this works.)
set.seed(84902)
hamby224_lime <- lime(x = hamby173and252_train %>% select(rf_features),
                      model = as_classifier(rtrees))

# Save the lime object
saveRDS(hamby224_lime, "../data/hamby224_lime.rds")
```

```{r}
# Apply the explain function from the lime package (with a seed set)
set.seed(84902)
hamby224_explain <- lime::explain(hamby224_test %>% select(rf_features), 
                                  hamby224_lime, 
                                  n_labels = 1, 
                                  n_features = 3)

# Save the explainer object
saveRDS(hamby224_explain, "../data/hamby224_explain.rds")
```

```{r echo = FALSE}
# Add a case variable to the test data
hamby224_test <- hamby224_test %>%
  mutate(case = as.character(1:dim(hamby224_test)[1])) %>%
  select(case, study:samesource)

# Obtain features used when fitting the rtrees random forest
rf_features <- rownames(rtrees$importance)

# Join the data and the explanations and edit and add additional variables
hamby224_test_explain <- full_join(hamby224_test, hamby224_explain, by = "case") %>%
  mutate(case = factor(case),
         set = factor(set),
         bullet1 = factor(bullet1),
         bullet2 = factor(bullet2),
         land1 = factor(land1),
         land2 = factor(land2)) %>%
  mutate(mypred = rep(predict(rtrees, hamby224_test %>% select(rf_features), 
                              type = "prob")[,2], each = 3)) %>%
  select(case:rfscore, mypred, samesource:prediction) %>%
  mutate(feature_desc = factor(feature_desc),
         feature = factor(feature)) %>%
  mutate(feature_number = readr::parse_number(feature_desc),
         strictly_less = FALSE)

@@

# Save the combined test and explain data
saveRDS(hamby224_test_explain, "../data/hamby224_test_explain.rds")
```

@@

```{r}
hamby224_test_explain %>%
  filter(set == "1") %>%
  select(case, bullet1:land2, samesource) %>%
  distinct() %>%
  ggplot(aes(x = land1, y = land2)) +
  geom_tile(aes(fill = samesource)) +
  facet_grid(bullet1 ~ bullet2) +
  theme_bw()
```

```{r, fig.width = 8, fig.height = 4}
lime::plot_explanations(hamby224_explain)
```

```{r}
hamby224_test_explain %>%
  filter(set == "1") %>%
  ggplot(aes(x = case, y = feature_desc)) + 
  geom_tile(aes(fill = feature_weight)) + 
  facet_grid(feature ~ label, space = "free", scale = "free") + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_fill_gradient2(midpoint = 0) +
  theme(legend.position = "bottom") +
  ylab("Feature Sets") 
```

```{r}
hamby224_test_explain %>%
  filter(set == "1", samesource == "FALSE") %>%
  ggplot(aes(x = case, y = feature_desc)) + 
  geom_tile(aes(fill = feature_weight)) + 
  facet_grid(feature ~ bullet1, space = "free", scale = "free") + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_fill_gradient2(midpoint = 0)
```

```{r}
hamby224_test_explain %>%
  filter(set == "1", samesource == "FALSE") %>%
  ggplot(aes(x = case, y = feature_desc)) + 
  geom_tile(aes(fill = feature_weight)) + 
  facet_grid(bullet1 ~ bullet2, space = "free", scale = "free") + 
  theme(axis.text.x = element_text(angle = 90)) +
  scale_fill_gradient2(midpoint = 0) +
  theme(legend.position = "bottom")
```

>>>>

# If the test data does not exist, clean the data, create the test data, and save it
# Otherwise, load in the test data
if(!file.exists("../data/hamby224_test.csv")) {

@@

  # Join the two cleaned Hamby 224 sets into one testing set
  hamby224_test <- suppressWarnings(bind_rows(hamby224_set1_cleaned,
                                              hamby224_set11_cleaned)) %>%
    mutate(case = as.character(1:length(study))) %>%
    select(case, study:samesource)

  # Save the test data as a .csv file
  write.csv(hamby224_test, "../data/hamby224_test.csv", row.names = FALSE)

} else {

  # Read in the test data
  hamby224_test <- read.csv("../data/hamby224_test.csv")

}
```

@@

```{r echo = FALSE}
# Apply the lime function if the file does not already exist
if(!file.exists("../data/hamby224_lime.rds")) {

  # Apply the lime function from the lime package (with a seed set)
  # Note that the as_classifier must be added since rtrees is from the
  # randomForest package and not fit using caret or one of the other 
  # available models specified in the lime package. Additionally, the
  # randomForest package must be loaded in order to run the functions
  # from the lime package. (I should check on exactly how this works.)
  set.seed(84902)
  hamby224_lime <- lime(x = hamby173and252_train %>% select(rf_features),
                        model = as_classifier(rtrees))

  # Save the lime object
  saveRDS(hamby224_lime, "../data/hamby224_lime.rds")

} else {

  # Load in the lime object
  hamby224_lime <- readRDS("../data/hamby224_lime.rds")

}
```

```{r}
# Apply explain function if the file does not already exist
if(!file.exists("../data/hamby224_explain.rds")) {

  # Apply the explain function from the lime package (with a seed set)
  set.seed(84902)
  hamby224_explain <- lime::explain(hamby224_test %>% select(rf_features), 
                                    hamby224_lime, 
                                    n_labels = 1, 
                                    n_features = 3)

  # Save the explainer object
  saveRDS(hamby224_explain, "../data/hamby224_explain.rds")

} else {

  # Load in the explain object
  hamby224_explain <- readRDS("../data/hamby224_explain.rds")

}
```

```{r echo = FALSE}
# Create the test_explain combined data if the file does not already exist
if(!file.exists("../data/hamby224_test_explain.rds")) {

  # Join the data and the explanations and edit and add additional variables
  hamby224_test_explain <- hamby224_test %>%
    mutate(case = as.character(case)) %>%
    full_join(hamby224_explain, by = "case") %>%
    mutate(case = factor(case),
           set = factor(set),
           bullet1 = factor(bullet1),
           bullet2 = factor(bullet2),
           land1 = factor(land1),
           land2 = factor(land2)) %>%
    mutate(mypred = rep(predict(rtrees, hamby224_test %>% select(rf_features), 
                                type = "prob")[,2], each = 3)) %>%
    select(case:rfscore, mypred, samesource:prediction) %>%
    mutate(feature_desc = factor(feature_desc),
           feature = factor(feature)) %>%
    mutate(feature_number = readr::parse_number(feature_desc),
           strictly_less = FALSE)

@@

  # Save the combined test and explain data
  saveRDS(hamby224_test_explain, "../data/hamby224_test_explain.rds")

} else {

  # Load in the data
  hamby224_test_explain <- readRDS("../data/hamby224_test_explain.rds")

}
```

@@
